---
title: "Origami Prediction"
author: "Five Find Outers"
output: html_document
---
```{r}
library(caret)
library(RWeka)
library(tidyverse)
library(GGally)
library(lubridate)
library(party)
library(rpart)
library(rpart.plot)
```


```{r}
drop_out <- read.csv("Vidya data 1 - Sheet2.csv")
```

```{r}
set.seed(1313)
drop_out <- drop_out %>% mutate(TC_ID = as.factor(TC_ID), Gender = as.factor(Gender), Country = as.factor(Country), Weekly_academics = as.factor(Weekly_academics), Weekly_extracurriculars = as.factor(Weekly_extracurriculars), Technology = as.factor(Technology), Internet_speed = as.factor(Internet_speed), Enjoy_AC = as.factor(Enjoy_AC), Relaxing = as.factor(Relaxing), Instruction = as.factor(Instruction), Class.format = as.factor(Class.format), Completed = factor(Completed, levels = c("No", "Yes")), Take.online.class = factor(Take.online.class, levels = c("No","Yes")))
```

```{r}
perc_75 <- (unique(drop_out$TC_ID) %>% length) * .75
perc_75 <- perc_75 %>% round
train_ids <- sample(unique(drop_out$TC_ID))[1:perc_75]
drop_out <- drop_out %>% mutate(type = ifelse(TC_ID %in% train_ids, "train", "test"))
drop_out_train <- drop_out %>% filter(type == "train") 
drop_out_test <- drop_out %>% filter(type == "test") 
```
Using the CART Trees Model

```{r}
library(caret)
TRAIN1 <- drop_out_train %>% select(-type)
TRAIN2 <- TRAIN1[,c(2:13)]
TEST1 <- drop_out_test[,c(2:13)]
#Removing the Type and TC_ID columns
```
Prediction1 - Predicting the Completion of the Origami Class
Training a basic model
```{r}
model1 <- rpart(Completed~., data = TRAIN2, control=rpart.control(minsplit=2, minbucket=1, cp=0.001))
```
```{r}
par(xpd = NA) # Avoid clipping the text in some device
plot(model1)
text(model1, digits = 3)
```
Testing the model
```{r}
predicted.classes <- model1 %>% 
  predict(TEST1, type = "class")
head(predicted.classes)
```
```{r}
mean(predicted.classes == TEST1$Completed)
```
The accuracy is 60%


Training a fitted model
```{r}

MySummary  <- function(data, lev = NULL, model = NULL){
  df <- defaultSummary(data, lev, model)
  tc <- twoClassSummary(data, lev, model)
  pr <- prSummary(data, lev, model)
  out <- c(df,tc,pr)
  out}

ctrl <- trainControl(method = "repeatedcv", #Tell caret to perform k-fold cross validation
                repeats = 3, #Tell caret to repeat each fold three times
                classProbs = TRUE, #Calculate class probabilities
                summaryFunction = MySummary)

cartFit <- train(Completed ~., #Define which variable to predict 
                data = TRAIN2, #Define the data set to train the model on
                trControl = ctrl,
                method = "rpart", #Define the model type
                metric = "Accuracy", #Final model choice is made according to sensitivity
                preProc = c("center", "scale")) #Center and scale the data to minimize the 
```

```{r}
#Check the results
cartFit$finalModel
cartFit
```
Accuracy is 71.66%

```{r}
library(rpart.plot)
rpart.plot(cartFit$finalModel)
print(cartFit)
```
Testing the fitted model
```{r setup, include=TRUE, cache = FALSE}
library(caret)
TEST1 <- drop_out_test[,c(2:13)] #Remove the student_id variable that we do not want to use in the model

#Generate prediction using previously trained model
cartClasses <- predict(cartFit,TEST1)

#Generate model statistics
confusionMatrix(data = cartClasses, as.factor(TEST1$Completed))
```
Accuracy is 60% as seen above, same as the basic model

Prediction2 - Predicting the choosing of an online Origami Class
Training a basic model
```{r}
TRAIN3 <- TRAIN1[,c(2:13)]
#Removing the Type and TC_ID columns
```
```{r}
model2 <- rpart(Take.online.class~., data = TRAIN2, control=rpart.control(minsplit=2, minbucket=1, cp=0.001))
```
```{r}
par(xpd = NA) # Avoid clipping the text in some device
plot(model2)
text(model2, digits = 3)
```
Testing the basic model
```{r}
predicted.classes <- model2 %>% 
  predict(TEST1, type = "class")
head(predicted.classes)
```
```{r}
mean(predicted.classes == TEST1$Take.online.class)
```
The accuracy is 80%

Training a fitted model
```{r}
MySummary  <- function(data, lev = NULL, model = NULL){
  df <- defaultSummary(data, lev, model)
  tc <- twoClassSummary(data, lev, model)
  pr <- prSummary(data, lev, model)
  out <- c(df,tc,pr)
  out}

ctrl <- trainControl(method = "repeatedcv", #Tell caret to perform k-fold cross validation
                repeats = 3, #Tell caret to repeat each fold three times
                classProbs = TRUE, #Calculate class probabilities
                summaryFunction = MySummary)

cartFit1 <- train(Take.online.class ~., #Define which variable to predict 
                data = TRAIN3, #Define the data set to train the model on
                trControl = ctrl,
                method = "rpart", #Define the model type
                metric = "Accuracy", #Final model choice is made according to sensitivity
                preProc = c("center", "scale")) #Center and scale the data to minimize the 
```

```{r}
#Check the results
cartFit1$finalModel
print(cartFit1)
rpart.plot(cartFit1$finalModel)
```
Testing the fitted model
```{r}
TEST1 <- drop_out_test[,c(2:13)]
cartClasses1 <- predict(cartFit1,TEST1)
```
```{r}
#Generate model statistics
confusionMatrix(data = cartClasses1, as.factor(TEST1$Take.online.class))
```
The accuracy is the same, at 80%



Conditional Inference trees
Prediction1 - Predicting the Completion of the Origami Class
```{r}
library(party)
```
Training a Conditional Inference Tree model
```{r}
condFit <- ctree(Completed~., TRAIN2)
plot(condFit, tp_args = list(beside = TRUE))
condFit
```
Testing the model
```{r}
condClasses <- predict(condFit, newdata = TEST1)

 #Generate model statistics
 confusionMatrix(data = condClasses, as.factor(TEST1$Completed))
```
The accuracy is 60%

Prediction2 - Predicting students taking online classes for Origami
Training a Conditional Inference Model
```{r}
condFit1 <- ctree(Take.online.class~., TRAIN2)
plot(condFit1, tp_args = list(beside = TRUE))
condFit1
```
Testing the model
```{r}
condClasses1 <- predict(condFit1, newdata = TEST1)

 #Generate model statistics
 confusionMatrix(data = condClasses1, as.factor(TEST1$Take.online.class))
```
The accuracy is noted at 80%

